---
title: "Practical Machine Learning - Coursera"
author: "Lakshay Kalbhor"
date: "24 October 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Practical Machine Learning Project

## Background

Utilizing gadgets such as Jawbone Up, Nike FuelBand, and Fitbit it is presently conceivable to gather a huge sum of information around individual movement generally cheaply. These sort of gadgets are portion of the evaluated self development - a gather of devotees who take estimations around themselves routinely to progress their wellbeing, to discover designs in their behavior. One thing that individuals frequently do is measure how much of a specific movement they do, but they seldom evaluate how well they do it. In this extend, your objective will be to utilize information from accelerometers on the belt, lower arm, arm, and dumbell of 6 members. They were inquired to perform barbell lifts accurately and inaccurately in 5 distinctive ways

## Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

# Analysis

Loading data:

```{r load data, warning=FALSE, message=FALSE, echo=TRUE}
training = read.csv("./pml-training.csv",na.strings=c("NA","#DIV/0!",""))
testing = read.csv("./pml-testing.csv",na.strings=c("NA","#DIV/0!",""))
# Data dimensions
dim(training)
dim(testing)
```

```{r first look, warning=FALSE, message=FALSE, eval= FALSE}
# First look at the data
head(training)
head(testing)
```

Cross-validation will be performed by spliting the preparing dataset into: 
1) A preparing dataset, containing 70% of the perceptions. The models for forecast will be built utilizing this dataset. 
2) A testing dataset, containing 30% of the perceptions. The precision of our expectation models will be assessed utilizing this dataset.

```{r cross-validation, warning=FALSE, message=FALSE, echo=TRUE}
# load packages
library(caret)
library(randomForest)
# Index for training dataset (70%) and testing dataset (30%) 
# from the pml-training data set
set.seed(12345)
inTrain = createDataPartition(y=training$classe,p=0.7, list=FALSE)
# training dataset
training.set = training[inTrain,]
# testing dataset
testing.set = training[-inTrain,]
```

Training and testing data consist of 160 variables. The choice of specific predictors is based on removing near zero variance predictors, with the nearZeroVar function, and also variables containing many NAs.

```{r clean data, warning=FALSE, message=FALSE, echo=TRUE}
# Remove near zero variance predictors
ind.nzv = nearZeroVar(x = training, saveMetrics = T)
# Remove variables with more than 50% NA values
ind.NA = !as.logical(apply(training, 2, function(x){ mean(is.na(x)) >= 0.5}))
# Cleaning data
ind2 = ind.NA*1 + (!ind.nzv$nzv)*1
ind3 = ind2 == 2
sum(ind3)
#View(data.frame(ind.NA, !ind.nzv$nzv, ind2, ind3))
training.set = training.set[,ind3]
testing.set = testing.set[, ind3]
training.set = training.set[, -1]
testing.set = testing.set[, -1]
testing = testing[,ind3]
testing = testing[,-1]
# Coerce the data into the same type in order to avoid
# "Matching Error" when calling random forest model, due to different levels in variables
for (i in 1:length(testing) ) {
  for(j in 1:length(training.set)) {
    if( length( grep(names(training.set[i]), names(testing)[j]) ) == 1)  {
      class(testing[j]) <- class(training.set[i])
    }      
  }      
}
# To get the same class between testing and training.set
testing = testing[,-ncol(testing)]
testing <- rbind(training.set[2, -58] , testing)
testing <- testing[-1,]
```

We are going utilize two approaches to form a forecast show for the values of classe variable. Firstly forecast with trees will be endeavored, utilizing the 'rpart' strategy and the caret bundle.

```{r prediction with trees, warning=FALSE, message=FALSE, echo=TRUE}
# Prediction with Trees
# Build model
set.seed(12345)
tree.fit = train(y = training.set$classe,
                 x = training.set[,-ncol(training.set)],
                 method = "rpart")
# Plot classification tree
rattle::fancyRpartPlot(
  tree.fit$finalModel
)
# Predictions with rpart model
pred.tree = predict(tree.fit, testing.set[,-ncol(testing.set)])
# Get results (Accuracy, etc.)
confusionMatrix(pred.tree, testing.set$classe)
```

Secondly a prediction model using random forest method will be created.

```{r random forest, warning=FALSE, message=FALSE, echo=TRUE}
# Prediction with Random Forest
# Build model
set.seed(12345)
rf.fit = randomForest(
  classe ~ .,
  data = training.set,
  ntree = 250)
# Plot the Random Forests model
plot(rf.fit)
# Predict with random forest model
pred2 = predict(
  rf.fit,
  testing.set[,-ncol(testing.set)]
)
# Get results (Accuracy, etc.)
confusionMatrix(pred2, testing.set$classe)
```

The exactness of the arbitrary timberland demonstrate is, as anticipated, much higher than the rpart demonstrate, over 0.99! Random Woodland show performed way better and constitutes the show of choice for foreseeing the 20 perceptions of the first pml-testing.csv dataset.

```{r pml-testing predictions, warning=FALSE, message=FALSE, echo=TRUE}
# Get predictions for the 20 observations of the original pml-testing.csv
pred.validation = predict(rf.fit, testing)
pred.validation
```

```{r saving results, warning=FALSE, message=FALSE, echo=TRUE, eval = FALSE}
# Saving predictions for testing dataset
testing$pred.classe = pred.validation
write.table(
  testing,
  file = "testing_with_predictions",
  quote = F
)
```

